
# Lab 1: Subset Selection Methods

> Best Subset Selection

```{r}
library(ISLR) 
attach(Hitters)
names(Hitters)
sum(is.na(Hitters$Salary)) 
```
Hence we see that Salary is missing for 59 players. The `na.omit()` function removes all of the rows that have missing values in any variable.

```{r}
Hitters=na.omit(Hitters) 
dim(Hitters)
sum(is.na(Hitters))
```

The `regsubsets()` function (part of the leaps library) performs __best subset selection__ by identifying the best model that contains a given number of predictors, where best is quantified using RSS. The syntax is the same as for lm(). The summary() command outputs the best set of variables for each model size.

```{r}
library(leaps)
regfit.full = regsubsets(Salary~., data=Hitters)
summary(regfit.full)
```
An asterisk indicates that a given variable is included in the corresponding model. For instance, this output indicates that the best two-variable model contains only `Hits` and `CRBI`. 

By default, `regsubsets()` only reports results up to the best eight-variable model. But the nvmax option can be used in order to return as many variables as are desired. Here we fit up to a 19-variable model.


```{r}
regfit.full=regsubsets(Salary~.,data=Hitters, nvmax=19) 

reg.summary=summary(regfit.full)
```
The summary() function also returns R2, RSS, adjusted R2, Cp, and BIC. We can examine these to try to select the best overall model.
```{r}
names(reg.summary)
```

For instance, we see that the R2 statistic increases from 32 %, when only one variable is included in the model, to almost 55%, when all variables are included. As expected, the R2 statistic increases monotonically as more variables are included.


```{r}
reg.summary$rsq
```
Plotting RSS, adjusted R2, Cp, and BIC for all of the models at once will help us decide which model to select. Note the type="l" option tells R to connect the plotted points with lines.

```{r}
reg.summary$rsq
```

__Plotting RSS, adjusted R2, Cp, and BIC for all of the models at once will help us decide which model to select. Note the type="l" option tells R to connect the plotted points with lines.__

```{r}
par(mfrow=c(2,2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS",
type="l")
plot(reg.summary$adjr2 ,xlab="Number of Variables ",
ylab="Adjusted RSq",type="l")

```
The `points()` command works like the `plot()` command, except that it puts points on a plot that has already been created, instead of creating a new plot. The `which.max()` function can be used to identify the location of the maximum point of a vector. We will now plot a red dot to indicate the model with the largest adjusted R2 statistic.

```{r}
which.max(reg.summary$adjr2)
```
```{r}
par(mfrow=c(2,2))
plot(reg.summary$rss ,xlab="Number of Variables ",ylab="RSS",
type="l")
plot(reg.summary$adjr2 ,xlab="Number of Variables ",
ylab="Adjusted RSq",type="l")
points(11,reg.summary$adjr2[11], col="red",cex=2,pch=20)
```
In a similar fashion we can plot the `Cp` and `BIC` statistics, and indicate the models with the smallest statistic using `which.min()`.

```{r}
plot(reg.summary$cp ,xlab="Number of Variables ",ylab="Cp", type='l')
which.min(reg.summary$cp )
points(10,reg.summary$cp [10],col="red",cex=2,pch=20)
which.min(reg.summary$bic )

plot(reg.summary$bic ,xlab="Number of Variables ",ylab="BIC",
type='l')
points(6,reg.summary$bic [6],col="red",cex=2,pch=20)
```
The regsubsets() function has a built-in plot() command which can be used to display the selected variables for the best model with a given number of predictors, ranked according to the BIC, Cp, adjusted R2, or AIC. To find out more about this function, type ?plot.regsubsets.

```{r}
plot(regfit.full,scale="r2")
plot(regfit.full,scale="adjr2") 
plot(regfit.full,scale="Cp")
plot(regfit.full,scale="bic")
```
The top row of each plot contains a black square for each variable selected according to the optimal model associated with that statistic. For instance, we see that several models share a BIC close to −150. However, the model with the lowest BIC is the six-variable model that contains only AtBat, Hits, Walks, CRBI, DivisionW, and PutOuts. We can use the coef() function to see the coefficient estimates associated with this model.
```{r}
coef(regfit.full ,6)
```

> Forward and Backward Stepwise Selection

We can also use the `regsubsets()` function to perform forward `stepwise` or `backward stepwise selection`, using the argument `method="forward"` or `method="backward"`.

```{r}
regfit.fwd=regsubsets(Salary~.,data=Hitters,nvmax=19, method ="forward")
summary(regfit.fwd)

regfit.bwd=regsubsets(Salary~.,data=Hitters ,nvmax=19, method ="backward")
summary(regfit.bwd)

```

For instance, we see that using forward stepwise selection, the best one- variable model contains only CRBI, and the best two-variable model ad- ditionally includes Hits. For this data, the best one-variable through six- variable models are each identical for best subset and forward selection. However, the best seven-variable models identified by forward stepwise se- lection, backward stepwise selection, and best subset selection are different.

```{r}
coef(regfit.full ,7)
coef(regfit.fwd ,7)
coef(regfit.bwd ,7)

```

> Choosing Among Models Using the Validation Set Approach and Cross-Validation

We just saw that it is possible to choose among a set of models of different sizes using Cp, BIC, and adjusted R2. We will now consider how to do this using the validation set and cross-validation approaches.

In order for these approaches to yield accurate estimates of the test error, we must use only the training observations to perform all aspects of model-fitting—including variable selection. Therefore, the determination of which model of a given size is best must be made using only the training observations. This point is subtle but important. __If the full data set is used to perform the best subset selection step, the validation set errors and cross-validation errors that we obtain will not be accurate estimates of the test error.__

```{r}
set.seed(1)
train=sample(c(TRUE,FALSE), nrow(Hitters),rep=TRUE)
test =(!train)
```
Now, we apply `regsubsets()` to the training set in order to perform best subset selection.
```{r}
regfit.best=regsubsets(Salary~.,data=Hitters[train,], nvmax =19)
```

Notice that we subset the Hitters dat frame directly in the call in order to access only the training subset of the data, using the expression `Hitters[train,]`. We now compute the validation set error for the best model of each model size. We first make a model matrix from the test data.

```{r}
test.mat=model.matrix(Salary~.,data=Hitters[test,])
```
The `model.matrix()` function is used in many regression packages for building an “X” matrix from data. Now we run a loop, and for each size i, we extract the coefficients from regfit.best for the best model of that size, multiply them into the appropriate columns of the test model matrix to form the predictions, and compute the test MSE.

```{r}
val.errors=rep(NA,19)
for(i in 1:19) {
  coefi=coef(regfit.best,id=i)
  pred=test.mat[,names(coefi)]%*%coefi
  val.errors[i]=mean((Hitters$Salary[test]-pred)^2)
}
val.errors
```
We find that the best model is i=7 variables

```{r}
which.min(val.errors)
```
```{r}
coef(regfit.best ,7)
```
We will create a function to capture this process

```{r}
predict.regsubsets =function (object, newdata, id,...){
  form=as.formula(object$call [[2]])
  mat=model.matrix(form,newdata)
  coefi=coef(object ,id=id)
  xvars=names(coefi)
  mat[,xvars]%*%coefi
}
```
We demonstrate how we use this function below, when we do cross-validation.

```{r}
regfit.best=regsubsets(Salary~.,data=Hitters ,nvmax=19)
coef(regfit.best,7)
```

```{r}
k=10
set.seed(1)
folds=sample(1:k,nrow(Hitters),replace=TRUE)
cv.errors=matrix(NA,k,19, dimnames=list(NULL, paste(1:19)))

for (j in 1:k) {
  best.fit=regsubsets(Salary~., data=Hitters[folds!=j,], nvmax=19)
  for(i in 1:19){
    pred=predict(best.fit,Hitters[folds==j,],id=i) 
    cv.errors[j,i]=mean(x = (Hitters$Salary[folds==j]-pred)^2)
  }
}
```

This has given us a 10×19 matrix, of which the (i, j)th element corresponds to the test MSE for the ith cross-validation fold for the best j-variable model. We use the `apply()` function to average over the columns of this matrix in order to obtain a vector for which the jth element is the cross- validation error for the j-variable model

```{r}
mean.cv.errors=apply(cv.errors ,2,mean)
mean.cv.errors

par(mfrow=c(1,1))
plot(mean.cv.errors ,type='b')
```

We see that cross-validation selects an 11-variable model. We now perform best subset selection on the full data set in order to obtain the 10-variable model.
```{r}
reg.best=regsubsets (Salary~.,data=Hitters , nvmax=19)
coef(reg.best,10)
```

# Lab 2: Ridge Regression and the Lasso

```{r}
x=model.matrix(Salary~.,Hitters)[,-1] 
y=Hitters$Salary
```

> Ridge regression

```{r}
library(glmnet)
grid=10^seq(10,-2,length=100)
ridge.mod=glmnet(x,y,alpha=0,lambda=grid)
```
By default the `glmnet()` function performs ridge regression for an automati- cally selected range of λ values. However, here we have chosen to implement the function over a grid of values ranging from $λ = 10^10$ to $λ = 10^−2$, essentially covering the full range of scenarios from the null model containing only the intercept, to the `least squares fit`. As we will see, we can also compute model fits for a particular value of `λ` that is not one of the original grid values. Note that by default, the` glmnet()` function standardizes the variables so that they are on the same scale. To turn off this default setting, use the argument `standardize=FALSE`.

Associated with each value of λ is a vector of ridge regression coefficients, stored in a matrix that can be accessed by coef(). In this case, it is a 20×100 matrix, with 20 rows (one for each predictor, plus an intercept) and 100 columns (one for each value of λ).

```{r}
dim(coef(ridge.mod))
```
We expect the coefficient estimates to be much smaller, in terms of $l2$ norm, when a large value of λ is used, as compared to when a small value of λ is used. These are the coefficients when λ = 11,498, along with their $l2$ norm
```{r}
ridge.mod$lambda [50]
coef(ridge.mod)[,50]
sqrt(sum(coef(ridge.mod)[-1,50]^2))
```
In contrast, here are the coefficients when λ = 705, along with their l2 norm. Note the much larger l2 norm of the coefficients associated with this smaller value of λ.
```{r}
ridge.mod$lambda [60]
coef(ridge.mod)[,60]
sqrt(sum(coef(ridge.mod)[-1,60]^2))
```
We can use the predict() function for a number of purposes. For instance, we can obtain the ridge regression coefficients for a new value of λ, say 50:
```{r}
predict(ridge.mod,s=50,type="coefficients")[1:20,]
```
We now split the samples into a training set and a test set in order to estimate the test error of ridge regression and the lasso. 

We first set a random seed so that the results obtained will be repro- ducible.
```{r}
set.seed(1)
train=sample(1:nrow(x), nrow(x)/2)
test=(-train)
y.test=y[test]
```
Next we fit a ridge regression model on the training set, and evaluate its MSE on the test set, using λ = 4. Note the use of the `predict()` function again. This time we get predictions for a test set, by replacing `type="coefficients"` with the `newx` argument.

```{r}
ridge.mod=glmnet(x[train,],y[train], alpha=0, lambda=grid, thresh=1e-12)
ridge.pred=predict(ridge.mod,s=4,newx=x[test,]) 
mean((ridge.pred-y.test)^2)
```
The test MSE is 142199.2. Note that if we had instead simply fit a model with just an intercept, we would have predicted each test observation using the mean of the training observations. In that case, we could compute the test set MSE like this:
```{r}
mean((mean(y[train])-y.test)^2)
```

We could also get the same result by fitting a ridge regression model with a very large value of λ. Note that 1e10 means 1010.

```{r}
ridge.pred=predict(ridge.mod,s=1e10,newx=x[test,]) 
mean((ridge.pred-y.test)^2)
```
So fitting a ridge regression model with λ = 4 leads to a much lower test MSE than fitting a model with just an intercept. We now check whether there is any benefit to performing ridge regression with λ = 4 instead of just performing least squares regression. Recall that least squares is simply ridge regression with λ = 0.

```{r}
ridge.pred=predict(ridge.mod,s=0, newx=x[test,]) 

mean((ridge.pred-y.test)^2)
lm(y~x, subset=train)
predict(ridge.mod,s=0,type="coefficients")[1:20,]
```
n general, if we want to fit a (unpenalized) least squares model, then we should use the lm() function, since that function provides more useful outputs, such as standard errors and p-values for the coefficients.

In general, instead of arbitrarily choosing λ = 4, it would be better to use cross-validation to choose the tuning parameter λ. We can do this using the built-in cross-validation function, cv.glmnet(). By default, the function performs ten-fold cross-validation, though this can be changed using the argument nfolds. Note that we set a random seed first so our results will be reproducible, since the choice of the cross-validation folds is random.

```{r}
set.seed(1)
cv.out = cv.glmnet(x[train,], y[train], alpha=0)
plot(cv.out)
bestlam = cv.out$lambda.min
bestlam
```

Therefore, we see that the value of λ that results in the smallest cross- validation error is 326. What is the test MSE associated with this value of λ?
```{r}
ridge.pred=predict(ridge.mod,s=bestlam ,newx=x[test,])
mean((ridge.pred-y.test)^2)
```
This represents a further improvement over the test MSE that we got using λ = 4. Finally, we refit our ridge regression model on the full data set, using the value of λ chosen by cross-validation, and examine the coefficient estimates.

```{r}
out=glmnet(x,y,alpha=0)
predict(out,type="coefficients",s=bestlam)[1:20,]
```
As expected, none of the coefficients are zero—ridge regression does not perform variable selection!

> The Lasso

We saw that ridge regression with a wise choice of λ can outperform least squares as well as the null model on the Hitters data set. We now ask whether the lasso can yield either a more accurate or a more interpretable model than ridge regression. In order to fit a lasso model, we once again use the glmnet() function; however, this time we use the argument alpha=1. Other than that change, we proceed just as we did in fitting a ridge model.

```{r}
lasso.mod=glmnet(x[train,],y[train],alpha=1,lambda=grid)
plot(lasso.mod)

```

We can see from the coefficient plot that depending on the choice of tuning parameter, some of the coefficients will be exactly equal to zero. We now perform cross-validation and compute the associated test error.

```{r}
set.seed(1)
cv.out=cv.glmnet(x[train ,],y[train],alpha=1)
plot(cv.out)
bestlam=cv.out$lambda.min

lasso.pred=predict(lasso.mod,s=bestlam ,newx=x[test,])
mean((lasso.pred-y.test)^2)
```

This is substantially lower than the test set MSE of the null model and of least squares, and similar to the test MSE of ridge regression with λ chosen by cross-validation.

However, the lasso has a substantial advantage over ridge regression in that the resulting coefficient estimates are sparse. Here we see that 12 of the 19 coefficient estimates are exactly zero. So the lasso model with λ chosen by cross-validation contains only eleven variables.
```{r}
out=glmnet(x,y,alpha=1,lambda=grid)
lasso.coef=predict(out,type="coefficients",s=bestlam)[1:20,]
lasso.coef

```

# Lab 3: PCR and PLS Regression

> Principal Components Regression

Principal components regression (PCR) can be performed using the pcr() function,whichispartoftheplslibrary.WenowapplyPCRtotheHitters pcr() data, in order to predict Salary. Again, ensure that the missing values have
been removed from the data, as described in Section 6.5.

```{r}
library(pls)
set.seed(2)
pcr.fit=pcr(Salary~., data=Hitters ,scale=TRUE, validation ="CV")
```

The syntax for the pcr() function is similar to that for lm(), with a few additional options. Setting scale=TRUE has the effect of standardizing each predictor, using (6.6), prior to generating the principal components, so that the scale on which each variable is measured will not have an effect. Setting validation="CV" causes pcr() to compute the ten-fold cross-validation error for each possible value of M , the number of principal components used. The resulting fit can be examined using summary().

```{r}
summary(pcr.fit)
```
The CV score is provided for each possible number of components, ranging from M = 0 onwards.

Note that pcr() reports the root mean squared error; in order to obtain the usual MSE, we must square this quantity. For instance, a root mean squared error of 352.8 corresponds to an MSE of 352.82 = 124,468.

One can also plot the cross-validation scores using the validationplot() function. Using val.type="MSEP" will cause the cross-validation MSE to be plot() plotted.

```{r}
validationplot(pcr.fit,val.type="MSEP")
```
We see that the smallest cross-validation error occurs when M = 16 com- ponents are used. This is barely fewer than M = 19, which amounts to simply performing least squares, because when all of the components are used in PCR no dimension reduction occurs. However, from the plot we also see that the cross-validation error is roughly the same when only one component is included in the model. This suggests that a model that uses just a small number of components might suffice.

The summary() function also provides the percentage of variance explained in the predictors and in the response using different numbers of compo- nents. This concept is discussed in greater detail in Chapter 10. Briefly, we can think of this as the amount of information about the predictors or the response that is captured using M principal components. For example, setting M = 1 only captures 38.31 % of all the variance, or information, in the predictors. In contrast, using M = 6 increases the value to 88.63 %. If we were to use all M = p = 19 components, this would increase to 100 %.
We now perform PCR on the training data and evaluate its test set performance.
```{r}
set.seed(1)
pcr.fit=pcr(Salary~., data=Hitters,subset=train,scale=TRUE,
validation ="CV")
validationplot(pcr.fit,val.type="MSEP")

```
Now we find that the lowest cross-validation error occurs when M = 5 component are used. We compute the test MSE as follows.

```{r}
pcr.pred=predict(pcr.fit,x[test,],ncomp=5) 
mean((pcr.pred-y.test)^2)
```

This test set MSE is competitive with the results obtained using ridge re- gression and the lasso. However, as a result of the way PCR is implemented, the final model is more difficult to interpret because it does not perform any kind of variable selection or even directly produce coefficient estimates.
Finally, we fit PCR on the full data set, using M = 7, the number of components identified by cross-validation.
```{r}
pcr.fit=pcr(y~x,scale=TRUE,ncomp=7)
summary(pcr.fit)
```

> Partial Least Squares

```{r}
set.seed(1)
pls.fit=plsr(Salary~., data=Hitters ,subset=train, scale=TRUE, validation ="CV") 

summary(pls.fit)
validationplot(pls.fit,val.type="MSEP")
   
```




